Moral Foundations of U.S. Political News Organizations

### 2023-07-21 manuscript_version/revision/manuscript_metapsych_revision.tex
here's the updates

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man,floatsintext]{apa6}
\title{Moral Foundations of U.S. Political News Organizations}
\author{William E. Padfield\textsuperscript{1}, Erin M. Buchanan, Ph.D.\textsuperscript{2}, \& Kayla N. Jordan, Ph.D.\textsuperscript{2}}
\date{}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Moral Foundations of U.S. Political News Organizations},
  pdfauthor={William E. Padfield1, Erin M. Buchanan, Ph.D.2, \& Kayla N. Jordan, Ph.D.2},
  pdflang={en-EN},
  pdfkeywords={politics, morality, psycholinguistics},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{MORAL NEWS}
\keywords{politics, morality, psycholinguistics}
\usepackage{csquotes}
\ifXeTeX
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[main=english]{babel}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\fi
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

@@

\abstract{
Partisan differences and diviseness have become an increasing hot topic in psychological research. Many theories have been proposed to explain these differences and divisions including Moral Foundations Theory. The current research seeks to use a linguistic measure of Moral Foundations, the Moral Foundations Dictionary (MFD), to test the theory in terms of predicted partisan differences. Through web scraping, we extracted articles from popular partisan news sources' websites, calculated MFD word frequencies, and identified words' respective valences. This process attempts to uncover news outlets' positive or negative endorsements of certain moral dimensions concomitant with a particular ideology. In Experiment 1, we gathered political articles from four sources. We were unable to reveal significant differences in moral endorsements, but we solidified the method to be employed in further research. In Experiment 2, we expanded their number of sources to 10 and analyzed articles that pertain to two specific topics: the 2018 confirmation hearings of U.S. Supreme Court Justice Brett Kavanaugh and the partial U.S. Government Shutdown of 2018-2019. Once again, no significant differences in moral endorsements were found. Together with past work, the results shed doubt on the validity of the MFD as a reliable measurement tool.
}

@@

Beyond the critiques of the theory itself, many have also provided strong critiques of the measurement using MFD.Conducting two close conceptual replication and six extension studies, Frimer (2020) found that the liberal-conservative differences found in the original study replicated in only 30\% of cases and the effect sizes were over 30 times smaller than reported in Graham et al. (2009). In a meta-analysis, the theorized differences were only found for authority and purity (e.g., conservatives used more authority and purity words). Loyalty was contrary to theoretical predictions (e.g., used more by liberals), and no differences were found for harm and fairness. Other studies have also found conflicting findings which fail to support the basic assertions of MFT. Exploring the political Twitterverse, Sterling and Jost (2018) found differences in the MFD depend on political sophistication. For example, harm words were used more by liberals only if they had low levels of political sophistication; at high levels of political sophistication, conservatives used more harm words. While the MFD has not been extensively used (at least in published research), studies which have use the MFD to test partisan differences have found mixed results in terms of supporting MFT (Clifford \& Jerit, 2013; Frimer, Tell, \& Haidt, 2015; Sagi \& Dehghani, 2014).

The most successful attempts to validate a linguistic measure of moral foundations have used more complex methods (compared to the original dictionary-based approach). These attempts generally fall into two categories: human annotations and semantic representations. Hopp, Fisher, Cornell, Huskey, and Weber (2021) represents the first approach where instead of relying on `experts,' a large sample of lay people were crowdsourced to manually annotate document for content relevant to each foundation. While resource-intensive, this method better captures how people may encounter and judge moral issues in everyday life leading to a more ecologically valid dictionary. Garten, Boghrati, Hoover, Johnson, and Dehghani (2016) represents the second categories which instead relies on data-driven semantic analysis. Using shallow neural net models like word2vec (Mikolov, Chen, Corrado, \& Dean, 2013), moral foundations were measured using semantic similarity vectors rather than simple word counts.

As a further conceptual replication of Frimer (2020), we test the liberal-conservative difference proposed by MFT using partisan media content and the MFD. We first test differences in a general news corpus compared liberal sources (National Public Radio {[}NPR{]} and the New York Times {[}NYT{]}) to conservative sources (Fox News and Breitbart). Second, we examined news about specific political events: Brett Kavaugh's U.S. Supreme Court nomination and the U.S. Government Shutdown in 2018-2019. In the second experiment, the news sources were expanded to cover NYT, NPR, Slate, Huffington Post, and Politico on the liberal side and Fox News, Breitbart, The Rush Limbaugh Show, The Blaze, and Sean Hannity on the conservative side. To address potential issues of measurement with the MFD, the MFD was combined with previous research by the current authors (see below) and weighted by valence to create weighted percentages to better specify endorsement. Like Frimer (2020), results were judged based on the direction of the liberal-conservative difference compared MFT predictions and the size of the effect compared to past studies.

\hypertarget{experiment-1}{%
\section{Experiment 1}\label{experiment-1}}

@@

For Experiment 1, we approached the study with the intention to answer a method question. That is, this portion of the current research was conducted in order to solidify the best method by which to analyze political news text under the MFT framework while also alleviating some of the aforementioned valence problem observed in the MFD. We hypothesized the news sources generally perceived as liberal leaning (\emph{NPR} and \emph{The New York Times}) would contain MFD words and valences indicating endorsements of the individualizing moral foundations (\emph{harm/care} and \emph{fairness/reciprocity}). Additionally, we hypothesized the two sources generally perceived to be conservative leaning (\emph{Fox News} and \emph{Breitbart}) would feature MFD words and valences indicating equal endorsement of all five foundations.

@@

Political articles were collected from the websites of four notable U.S. news sources, a process known as web scraping. The sources were \emph{NYT}, \emph{NPR}, \emph{Fox News}, and \emph{Breitbart}. They were selected for their widespread recognition and the fact political partisans have strong preferences for some sources over others. We determined the political lean of each source by referencing Mitchell, Matsa, Gottfried, and Kiley (2014)'s article demonstrating the self-reported ideological consistency represented by the consumers of several news sources. In general, the \emph{NYT} and \emph{NPR} are preferred by consumers reporting a liberal bias or lean. In contrast, \emph{Fox News} and \emph{Breitbart} are believed to have a conservative bias or lean. Mitchell et al. (2014)'s article presented political ideology as a scale ranging from ``consistently liberal'' to ``consistently conservative.'' In between these extremes lie more moderate positions, including ``mostly liberal,'' ``mixed,'' and ``mostly conservative.'' Owing to the lower number of sources analyzed herein, we elected to categorize the sources as either ``liberal'' and ``conservative'' in order to form a basis for comparison.

@@

The source article words were compiled into a dataset where they were matched up with their counterparts in the MFD along with their valence and a percentage of their occurrence. Therefore, for each article, the percentage of the number of \emph{harm/care} words occurring in the articles were calculated, and this process was repeated for each of the foundations. Words' percent occurrence were multiplied by their \emph{z}-scored valence. Valences were \emph{z}-scored in order to eliminate any ambiguity regarding the direction of the valence. Positive values indicate positive valence, and negative values indicate negative valence. Words were categorized in accordance to their MFD affiliation, creating a weighted sum for each moral foundation.

@@

\caption{\label{tab:exp1-source-descriptives}Experiment 1 - Descriptive Statistics by Source}

@@

\begin{tabular}{lcccccccccc}
\toprule
Source & $M_V$ & $SD_V$ & $N_{Article}$ & $N_{Words}$ & $M_T$ & $SD_T$ & $M_{Ty}$ & $SD_{Ty}$ & $M_{FK}$ & $SD_{FK}$\\
\midrule
NPR & 0.28 & 0.23 & 695 & 302977 & 435.94 & 642.63 & 191.96 & 192.28 & 14.00 & 3.93\\
New York Times & 0.30 & 0.13 & 406 & 452579 & 1114.73 & 511.86 & 454.27 & 154.58 & 16.42 & 3.36\\
Breitbart & 0.29 & 0.18 & 1437 & 722022 & 502.45 & 347.90 & 243.35 & 120.75 & 18.57 & 7.89\\
Fox News & 0.29 & 0.17 & 503 & 296779 & 590.02 & 528.60 & 283.56 & 189.00 & 17.25 & 7.21\\
\bottomrule
\addlinespace
\end{tabular}

@@

To analyze if news sources adhered to differences in word use based on their target audience, we utilized a multilevel model (MLM) to analyze the data. MLM is a regression technique that allows one to control for the repeated measurement and nested structured of the data, which creates correlated error (Gelman, 2006). Using the \emph{nlme} library in \emph{R} (Pinheiro, Bates, Debroy, Sarkar, \& Team, 2017), each foundation's weighted percentage was predicated here by the political lean of the news source, using the individual news sources as a random intercept to control for the structure of the data.

@@

\caption{\label{tab:exp1-table}Experiment 1 Results - Multilevel Model}

@@

\begin{tabular}{lccccccc}
\toprule
Foundation & $M_C$ & $SD_C$ & $M_L$ & $SD_L$ & $t$ & $p$ & $d$\\
\midrule
Harm/Care & 0.50 & 2.21 & 0.49 & 2.21 & -0.21 & .851 & 0.01\\
Fairness/Reciprocity & 1.13 & 1.38 & 1.11 & 1.38 & -0.42 & .713 & 0.02\\
Ingroup/Loyalty & 1.28 & 1.63 & 1.34 & 1.63 & 0.30 & .793 & -0.04\\
Authority/Respect & 0.72 & 1.62 & 1.06 & 1.62 & 3.16 & .087 & -0.20\\
Purity/Sanctity & 1.11 & 1.48 & 1.27 & 1.48 & 2.38 & .140 & -0.09\\
\bottomrule
\addlinespace
\end{tabular}

@@

The results obtained in Experiment 1 fail to support the differences predicted by MFT. First, differences between liberal and conservative news sources failed to reach statistical or practical (effect size) significance for any foundations. Second, looking purely at the direction of the differences, the results run completely contrary to the differences expected by MFT. Conservative sources scored higher on concern for harm and fairness while liberal sources scored higher on concern for loyalty, authority, and purity. Overall, the results lend strong support to the case against the MFD and weakens the case generally for MFT.

However, one possible limitation of the current data is the generalness of the corpus. The selection of the broad and amorphous topic of ``political news'' may have led to the scraping of large numbers of articles with little to no moral-centric content. To address this limitation, two changes that were subsequently employed in Experiment 2. First, we elected to include more news sources for web scraping and analysis in addition to the four used in Experiment 1. Second, we chose to focus data collection efforts on two heavily moralized events in the Trump administration: (1) the nomination and confirmation of Justice Brett Kavanaugh to the U.S. Supreme Court and the U.S. government shutdown in December 2018 through January 2019 over disagreements about funding a U.S.-Mexico border wall. Hence, Experiment 2 is the best test of the MFT and the MFD in the context of partisan news.

\hypertarget{experiment-2}{%
\section{Experiment 2}\label{experiment-2}}

@@

As in Experiment 1, the text was scanned with \emph{ngram} (Schmidt et al., 2017). Again, blank articles, text from the Disqus system, and duplicate articles were removed. The text was processed and stemmed in order to convert to a usable form for further analysis (Feinerer \& Hornik, 2017). Each individual word was reduced to its stem (i.e., \emph{diseased} was stemmed to \emph{diseas}). Once again, the same procedure was applied to the MFD words and the words in the Warriner et al. (2013) dataset. We obtained the words' percent occurrence in the text. Percents were multiplied by \emph{z}-scored valence and categorized into their proper MFD category.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man,floatsintext]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

\makeatletter
\usepackage{etoolbox}
\patchcmd{\maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{politics, morality, psycholinguistics}
\usepackage{csquotes}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Moral Foundations of U.S. Political News Organizations},
  pdfauthor={William E. Padfield1, Erin M. Buchanan, Ph.D.2, \& Kayden N. Jordan, Ph.D.2},
  pdflang={en-EN},
  pdfkeywords={politics, morality, psycholinguistics},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Moral Foundations of U.S. Political News Organizations}
\author{William E. Padfield\textsuperscript{1}, Erin M. Buchanan, Ph.D.\textsuperscript{2}, \& Kayden N. Jordan, Ph.D.\textsuperscript{2}}
\date{}

\shorttitle{MORAL NEWS}

@@

\abstract{%
Partisan differences and diviseness have become an increasing hot topic in psychological research. Many theories have been proposed to explain these differences and divisions including Moral Foundations Theory. The current research seeks to use a linguistic measure of Moral Foundations, the Moral Foundations Dictionary (MFD), to test the theory in terms of predicted partisan differences. Through web scraping, we extracted articles from popular partisan news sources' websites, calculated MFD word frequencies, and identified words' respective valences. This process attempts to uncover news outlets' positive or negative endorsements of certain moral dimensions concomitant with a particular ideology. In Study 1, we gathered political articles from four sources. We were unable to reveal significant differences in moral endorsements, but we solidified the method to be employed in further research. In Study 2, we expanded their number of sources to 10 and analyzed articles that pertain to two specific topics: the 2018 confirmation hearings of U.S. Supreme Court Justice Brett Kavanaugh and the partial U.S. Government Shutdown of 2018-2019. Once again, no significant differences in moral endorsements were found. Together with past work, the results shed doubt on the validity of the MFD as a reliable measurement tool.
}

@@

Beyond the critiques of the theory itself, many have also provided strong critiques of the measurement using MFD. Conducting two close conceptual replication studies using religious sermons and six extension studies using U.S. Congressional speeches, party platforms, State of the Union addresses, media outlets, and political organization texts, Frimer (2020) found that the liberal-conservative differences found in the original study replicated in only 30\% of cases and the effect sizes were over 30 times smaller than reported in Graham et al. (2009). In a meta-analysis, the theorized differences were only found for authority and purity (e.g., conservatives used more authority and purity words). Loyalty was contrary to theoretical predictions (e.g., used more by liberals), and no differences were found for harm and fairness. Additionally, Frimer (2020) tested the theory under the most forgiving conditions using not only the original MFD but also the revised MFD2 as well using log transformed scores as well as raw scores. Even under the best circumstances (log transformed MFD2 scores), only 2 of the foundations showed theorized differences but with very small effects. Other studies have also found conflicting findings which fail to support the basic assertions of MFT. Exploring the political Twitterverse, Sterling and Jost (2018) found differences in the MFD depend on political sophistication. For example, harm words were used more by liberals only if they had low levels of political sophistication; at high levels of political sophistication, conservatives used more harm words. While the MFD has not been extensively used (at least in published research), studies which have use the MFD to test partisan differences have found mixed results in terms of supporting MFT (Clifford \& Jerit, 2013; Frimer, Tell, \& Haidt, 2015; Sagi \& Dehghani, 2014). Like Frimer (2020), we attempt to give the MFD the best chance for success by incorporating a weighting method to address the skewedness and sparsity of the linguistic data.

The most successful attempts to validate a linguistic measure of moral foundations have used more complex methods (compared to the original dictionary-based approach). These attempts generally fall into two categories: human annotations and semantic representations. Hopp, Fisher, Cornell, Huskey, and Weber (2021) represents the first approach where instead of relying on `experts', a large sample of lay people were crowdsourced to manually annotate document for content relevant to each foundation. While resource-intensive, this method better captures how people may encounter and judge moral issues in everyday life leading to a more ecologically valid dictionary. Garten, Boghrati, Hoover, Johnson, and Dehghani (2016) represents the second categories which instead relies on data-driven semantic analysis. Using shallow neural net models like word2vec (Mikolov, Chen, Corrado, \& Dean, 2013), moral foundations were measured using semantic similarity vectors rather than simple word counts.

As a further conceptual replication of Frimer (2020), we test the liberal-conservative difference proposed by MFT using partisan media content and the MFD. We first test differences in a general news corpus compared liberal sources (National Public Radio {[}NPR{]} and the New York Times {[}NYT{]}) to conservative sources (Fox News and Breitbart). Second, we examined news about specific political events: Brett Kavanaugh's U.S. Supreme Court nomination and the U.S. Government Shutdown in 2018-2019. In the second study, the news sources were expanded to cover NYT, NPR, Slate, Huffington Post, and Politico on the liberal side and Fox News, Breitbart, The Rush Limbaugh Show, The Blaze, and Sean Hannity on the conservative side. To address potential issues of measurement with the MFD, the MFD was combined with previous research by the current authors (see below) and weighted by valence to create weighted percentages to better specify endorsement. Like Frimer (2020), results were judged based on the direction of the liberal-conservative difference compared MFT predictions and the size of the effect compared to past studies.

\hypertarget{study-1}{%
\section{Study 1}\label{study-1}}

@@

For the first study, we approached the study with the intention to answer a method question. That is, this portion of the current research was conducted in order to solidify the best method by which to analyze political news text under the MFT framework while also alleviating some of the aforementioned valence problem observed in the MFD. We hypothesized the news sources generally perceived as liberal leaning (\emph{NPR} and \emph{The New York Times}) would contain MFD words and valences indicating endorsements of the individualizing moral foundations (\emph{harm/care} and \emph{fairness/reciprocity}). Additionally, we hypothesized the two sources generally perceived to be conservative leaning (\emph{Fox News} and \emph{Breitbart}) would feature MFD words and valences indicating equal endorsement of all five foundations.

@@

Political articles were collected from the websites of four notable U.S. news sources, a process known as web scraping. The sources were \emph{NYT}, \emph{NPR}, \emph{Fox News}, and \emph{Breitbart}. They were selected for their widespread recognition and the fact political partisans have strong preferences for some sources over others. We determined the political lean of each source by referencing Mitchell, Matsa, Gottfried, and Kiley (2014)'s article demonstrating the self-reported ideological consistency represented by the consumers of several news sources. In general, the \emph{NYT} and \emph{NPR} are preferred by consumers reporting a liberal bias or lean. In contrast, \emph{Fox News} and \emph{Breitbart} are believed to have a conservative bias or lean. Mitchell et al. (2014)'s article presented political ideology as a scale ranging from ``consistently liberal'' to ``consistently conservative.'' In between these extremes lie more moderate positions, including ``mostly liberal,'' ``mixed,'' and ``mostly conservative.'' Owing to the lower number of sources analyzed herein, we elected to categorize the sources as either ``liberal'' and ``conservative'' in order to form a basis for comparison. At the time of our data collection, these sources were also selected because they were open (i.e., no subscription required to access articles), and their websites were designed in a way that made webscraping possible.

@@

A complete list of the words for each moral foundation area was created from previous experiments. The total number of times each of those words was used in an article was calculated. This information was merged with the valence values from the Warriner et al. (2013) dataset. A weighted score was calculated by multiplying the frequency of occurrence within a document times the \emph{z}-scored valence values and divided by the word count for that article. The final score was multiplied by 100 to represented a weighted percentage. Finally, the weighted scores were summed within in each article to get a total weighted percentage of the use of moral area specific words by article. Valences were \emph{z}-scored in order to eliminate any ambiguity regarding the direction of the valence. Positive values indicate positive valence, and negative values indicate negative valence. Words were categorized in accordance to their MFD affiliation, creating a weighted sum for each moral foundation. See Table \ref{tab:example-weight} for an example.

\caption{\label{tab:example-weight}Example of Weighted Coding for Harm Words}

\begin{tabular}{llcccc}
\toprule
Source & Concept & Frequency & Total Words & Valence & Weighted Score\\
\midrule
Breitbart & abus & 14.00 & 846 & -2.77 & -4.59\\
Breitbart & cruel & 2.00 & 574 & -1.83 & -0.64\\
NY Times & unit & 19.00 & 1341 & 0.18 & 0.25\\
Breitbart & first & 18.00 & 1135 & 1.78 & 2.82\\
Fox News & comfort & 12.00 & 1053 & 1.71 & 1.95\\
NPR & attent & 6.00 & 1336 & 1.27 & 0.57\\
Breitbart & bruis & 3.00 & 290 & -1.43 & -1.48\\
Breitbart & accid & 3.00 & 196 & -1.97 & -3.02\\
Breitbart & cut & 11.00 & 1036 & -0.91 & -0.97\\
Breitbart & child & 8.00 & 622 & 1.68 & 2.16\\
NPR & injuri & 4.00 & 968 & -2.15 & -0.89\\
NY Times & drug & 32.00 & 1004 & -0.75 & -2.38\\
Breitbart & beat & 5.00 & 298 & -0.54 & -0.90\\
Fox News & weapon & 17.00 & 1175 & -0.87 & -1.26\\
Breitbart & speci & 3.00 & 390 & 0.15 & 0.12\\
\bottomrule
\addlinespace
\end{tabular}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} }
\end{tablenotes}

@@

\caption{\label{tab:exp1-source-descriptives}Study 1 - Descriptive Statistics by Source}

@@

\begin{tabular}{lcccccccccc}
\toprule
Source & $M_V$ & $SD_V$ & $N_{Article}$ & $N_{Words}$ & $M_T$ & $SD_T$ & $M_{Ty}$ & $SD_{Ty}$ & $M_{FK}$ & $SD_{FK}$\\
\midrule
NPR & 0.28 & 0.23 & 695 & 302977 & 435.94 & 642.63 & 191.96 & 192.28 & 14.00 & 3.93\\
New York Times & 0.30 & 0.13 & 406 & 452579 & 1114.73 & 511.86 & 454.27 & 154.58 & 16.44 & 3.36\\
Breitbart & 0.29 & 0.18 & 1437 & 722022 & 502.45 & 347.90 & 243.35 & 120.75 & 18.57 & 7.89\\
Fox News & 0.29 & 0.17 & 503 & 296779 & 590.02 & 528.60 & 283.56 & 189.00 & 17.25 & 7.21\\
\bottomrule
\addlinespace
\end{tabular}

@@

To analyze if news sources adhered to differences in word use based on their target audience, we utilized a multilevel model (MLM) to analyze the data. MLM is a regression technique that allows one to control for the repeated measurement and nested structured of the data, which creates correlated error (Gelman, 2006). Using the \emph{nlme} library in \emph{R} (Pinheiro, Bates, Debroy, Sarkar, \& Team, 2017), each foundation's weighted percentage was predicated here by the political lean of the news source, using the individual news sources as a random intercept to control for the structure of the data. Therefore, the models were calculated by: Weighted Score \textasciitilde{} Political Lean + (\textasciitilde1\textbar Source).

@@

\caption{\label{tab:exp1-table}Study 1 Results - Multilevel Model}

@@

\begin{tabular}{lcccccccl}
\toprule
Foundation & $M_C$ & $SD_C$ & $M_L$ & $SD_L$ & $t$ & $p$ & $d$ & ICC\\
\midrule
Harm/Care & 0.50 & 2.21 & 0.49 & 2.21 & -0.21 & .851 & 0.01 & .851\\
Fairness/Reciprocity & 1.13 & 1.38 & 1.11 & 1.38 & -0.42 & .713 & 0.02 & .713\\
Ingroup/Loyalty & 1.28 & 1.63 & 1.34 & 1.63 & 0.30 & .793 & -0.04 & .793\\
Authority/Respect & 0.72 & 1.62 & 1.06 & 1.62 & 3.16 & .087 & -0.20 & .087\\
Purity/Sanctity & 1.11 & 1.48 & 1.27 & 1.48 & 2.38 & .140 & -0.09 & .140\\
\bottomrule
\addlinespace
\end{tabular}

@@

The results obtained in Study 1 fail to support the differences predicted by MFT. First, differences between liberal and conservative news sources failed to reach statistical or practical (effect size) significance for any foundations. Second, looking purely at the direction of the differences, the results run completely contrary to the differences expected by MFT. Conservative sources scored higher on concern for harm and fairness while liberal sources scored higher on concern for loyalty, authority, and purity. Overall, the results lend strong support to the case against the MFD and weakens the case generally for MFT.

However, one possible limitation of the current data is the generalness of the corpus. The selection of the broad and amorphous topic of ``political news'' may have led to the scraping of large numbers of articles with little to no moral-centric content. To address this limitation, two changes that were subsequently employed in Study 2. First, we elected to include more news sources for web scraping and analysis in addition to the four used in Study 1. Second, we chose to focus data collection efforts on two heavily moralized events in the Trump administration: (1) the nomination and confirmation of Justice Brett Kavanaugh to the U.S. Supreme Court and the U.S. government shutdown in December 2018 through January 2019 over disagreements about funding a U.S.-Mexico border wall. Hence, Study 2 is the best test of the MFT and the MFD in the context of partisan news.

\hypertarget{study-2}{%
\section{Study 2}\label{study-2}}

@@

As in Study 1, the text was scanned with \emph{ngram} (Schmidt et al., 2017). Again, blank articles, text from the Disqus system, and duplicate articles were removed. The text was processed and stemmed in order to convert to a usable form for further analysis (Feinerer \& Hornik, 2017). Each individual word was reduced to its stem (i.e., \emph{diseased} was stemmed to \emph{diseas}). Once again, the same procedure was applied to the MFD words and the words in the Warriner et al. (2013) dataset. We obtained the words' percent occurrence in the text. Percents were multiplied by \emph{z}-scored valence and categorized into their proper MFD category.
